---
title: "UMD Fees"
---

Follow along with the demo to learn how to calculate percentage change using R and the Tidyverse. You should get started by loading the tidyverse in a new R block.


### Task 1: Load the tidyverse


```{r}
library(tidyverse)
```



### Task 2: Load the UMD Fees dataset from https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/umd_fees.csv

```{r}
umd_fees <- read_csv("https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/umd_fees.csv")
```

### Task 3: Calculate the percentage change in each category between Fall 2024 and Fall 2021 using `mutate()`

```{r}
umd_fees_pct <- umd_fees |> mutate(pct_change = (`Fall 2024`- `Fall 2021`)/`Fall 2021` * 100)
```


```{r}
umd_fees_pct_2425 <- umd_fees |> mutate(pct_change = (`Fall 2025`- `Fall 2024`)/`Fall 2024` * 100)
```



### Task 4: Describe our results

Write a few sentences about what you found in the data. What's the story here?

Based on what I found in the data, I discovered that student sustainability had one of the largest percent changes along side the Student Counseling Center which had 0 funding in 2021. One story I would be interested in pursuing is what is student sustainability? Based on the stats, it needed an almost 150% increase in funding. Why? What does student sustainability involve? 




```{r}
comp <- read_csv("https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/class_composition.csv")

comp |>
  summarize(
    mean_first = mean(first_year),
    mean_soph = mean(sophomore),
    mean_junior = mean(junior),
    mean_senior = mean(senior),
    mean_grad = mean(grad)
  )
```


When you are done, save your work, switch to GitHub Desktop, then add, commit and push your changes to GitHub and submit the URL of the notebook in ELMS.
