---
title: "final_project"
output: html_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
library(tidyverse)
library(readxl)
url <- "https://opendoorsdata.org/wp-content/uploads/2025/11/OD25_US-Study-Abroad_Tables.xlsx"
download.file(url, destfile = "open_doors.xlsx", mode = "wb")
excel_sheets("open_doors.xlsx")
blackstudy_abroad <- read_excel("open_doors.xlsx", sheet = 2, skip = 3)
colnames(blackstudy_abroad) <- c("year", "total", "annual_change")
```
I learned (through Google and prayers lol) how to load an excel sheet/file. This data contains the study abroad rates over the years for Black students. 

```{r, clean_up}
blackstudy_abroad_clean <- blackstudy_abroad %>%
  mutate(
    year = as.character(year),
    total = as.numeric(total),
    annual_change = as.numeric(annual_change)
  ) %>%
  filter(
    !is.na(year),
    year != "Year",
    !is.na(total)
  )
blackstudy_abroad_clean <- blackstudy_abroad_clean %>%
  mutate(year = factor(year, levels = year))
```

```{r, graph_data}
ggplot(blackstudy_abroad_clean, aes(x = year, y = total, group = 1)) +
  geom_line() +
  geom_point() +
  theme_minimal() +
  labs(
    title = "U.S. Study Abroad Totals, 1985/86 - 2023/24",
    x = "Year",
    y = "Total Students"
  ) +
  scale_y_continuous(labels = scales::comma)
```

```{r, black_enrollment_data}
black_enrollment <- read_csv("/Users/savannahgrooms/Documents/table.csv", skip = 2)
```

So, after with some playing around with my topic and data, I settled on a data set by https://opendoorsdata.org/data/us-study-abroad/student-profile/, which provides national data on study abroad participation by race and year. I loaded the data, cleaned the data and was able to plot it on a ggplot. The data ranges from years 1985/86 - 2023/24. After taking your advice, I also found a dataframe from https://nces.ed.gov/programs/digest/d24/tables/dt24_306.10.asp on the "Total fall enrollment in degree-granting postsecondary institutions, by level of enrollment, sex, attendance status, and race/ethnicity". This data ranges from 1976-2023. I plan on working to clean up this table where it only has the Black race undergraduate students and total undergraduate students so I would be able to calculate a percent change. My biggest question is once I have these two data frames, how would I work to compare them?
