---
title: "HW3 WH Salaries"
---

In this assignment, we will be working with a dataset containing information about the salaries of White House staffers as of July 2025.

### Task 1: Load the Tidyverse

You will need to load the tidyverse library

```{r}
library(tidyverse)
```

Then you'll load the data from the web. The URL for the data is: https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/white_house_salaries_2025.csv. You will create a dataframe called `wh_salaries` that contains the data from this URL.

### Task 2: Load the Data and Describe It

```{r}
wh_salaries <- read_csv("https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/white_house_salaries_2025.csv")
```
### Task 3: Describe the Data

You should examine that dataframe carefully by clicking on it in the Environment tab in RStudio. Write a couple of sentences describing the categories of incidents. From a reporting standpoint, what stands out to you about this data? Are there any salary values that will cause problems for analysis?

While reviewing the data, I realized that most White Hosue employees make close to $100,000. The "staff/press assistants" had the lowest paying jobs within the White House. Their salaries is around 65,000. Even stranger, some assistants to the president had a salary of 0 while others with a similar title could make upwards to $120,000. The journalist in me had to research Karoline Leavitt's salary which was less than I imagined at around $195,000. There are a few values which may mess up my data including the various $0 salaries and the outlier of $225,700 to an advisor. 
### Task 4: Calculate the Mean and Median of the Salaries

Write code to calculate the mean and median of the salaries in the dataset using the `summarize()` function. You don't need to make a new dataframe, just display the results.

```{r}
wh_salaries |>
  summarize(
    mean_salary= mean(SALARY),
    median_salary = median(SALARY)
  )

```

### Task 5: Add in the minimum and maximum salaries

Take the code you just wrote and add in the minimum and maximum salaries (the functions for that are `min()` and `max()`). Your results should show four columns: mean_salary, median_salary, min_salary, max_salary.

```{r}
wh_salaries |>
  summarize(
    mean_salary= mean(SALARY),
    median_salary = median(SALARY),
    min_salary= min(SALARY),
    max_salary= max(SALARY)   
  )
```

### Task 6: Describe the Results

Write a couple of sentences describing the results of your analysis. What do the mean and median salaries tell you about the data? What do the minimum and maximum salaries tell you? Are there any outliers in the data?

The average salary is a bit higher than the median salary. With the outliers of $0 salaries and the $225,700 salary, I believe the median salary is more representative of data. The minimum and maximum salaries tell me that there is a very large range. These values affect how spread out the data appears. 

### Task 7: Story Idea

Based on the data, what is your best story idea? What is the next step you would take to develop that story idea?

Based on the data, the best story to pursue is a profile on JACALYNNE KLOPP, the top earning employee at the White House. Her title is labeled under "advisor". The next steps to develop this story is to research the type of advising she does. Why is this amount of money invested into her job? Interview other employees (possibly off-record) to see how they feel about the amount of money she makes compared to other key employees. 
