---
title: "HW4 MD CityCrime"
---

In this assignment, we will be working with a dataset of total crimes reported by Maryland cities for two consecutive years. You'll load the data, calculate the violent and property crime rates for each city for both years, and then calculate the percentage change in the crime rates between the two years. You'll also write a short story based on your findings.

### Task 1: Load the Tidyverse and Janitor

You will need to load the tidyverse library and the janitor library

```{r}
library(tidyverse)
```

Then you'll load the data from the web. The URL for the data is: https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/md_cities_crime_2019_2020.csv. You will create a dataframe called `md_cities` that contains the data from this URL.

### Task 2: Load the Data

```{r}
mdcities <-read_csv("https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/md_cities_crime_2019_2020.csv")
```
### Task 3: Calculating Rates

Calculate violent and property crime rates for each city for both years using `mutate()`. You'll create four new columns: `violent_rate_2019`, `property_rate_2019`, `violent_rate_2020`, and `property_rate_2020`. You'll need to figure out the proper unit rate, given that there are varying populations. There's no perfect choice; choose the one that works best for you. Save the results to a new dataframe called `md_cities_rates`.

```{r}
md_cities_rates <- mdcities |> 
  mutate (violent_rate_2019 = violent2019/pop2020 * 10, 
         property_rate_2019 = property2019/pop2020 * 10,
         violent_rate_2020 = violent2020/pop2020 * 10,
         property_rate_2020 = property2020/pop2020 * 10
         )
         
```

### Task 4: Describe the Rates

Examine the new dataframe, `md_cities_rates`, by clicking on it in the Environment tab in RStudio and sorting by the rate columns. Write a couple of sentences about the results, comparing and contrasting the cities with the highest rates (both violent and property) in 2019 and 2020. Do you see any patterns? What does this tell you about crime in Maryland cities? What's the story here?

One thing that really stood out to me was Baltimore had by far the highest violent crime rate in both years. In terms of property crime rates,  places like Ocean City (which I assume is because of its high tourist gaze). Smaller towns occasionally appear high on the list due to their low population. 
I think that crime is pretty much so high in more city central places. 


### Task 5: Calculate Percentage Change

Let's add two percentage change columns to your dataframe: one for violent crime rates and one for property crime rates. You'll calculate the percentage change between 2019 and 2020 for each city. Save the results to a new dataframe called `md_cities_change`.

```{r}
md_cities_change <- md_cities_rates |> 
 mutate(
    violent_change = (violent2020 - violent2019) / violent2019 * 100,
    property_change = (property2020 - property2019) / property2019 * 100
  )
```

### Task 7: Describe Percentage Changes

Look at the `md_cities_change` dataframe by clicking on it in the Environment tab in RStudio. Write a couple of sentences describing the percentage change results. What's more interesting here, the patterns or the outliers? What's the story in the data?

While observing the data I noticed that Baltimore, the largest city, saw double digit decreases in both violent and property crime from 2019 and 2020 which I can conclude means a overall decline in reported crime rates.On the other hand, smaller cities seemed to have a spike in violent and property crime such as Accident city and cottage city a mid-sized cities. Some smaller cities were labeled as inf because crime counts were very low to begin with, so even one additional incident dramatically changed the percentage.
I don't think the outliers are quite the story here but more so the pattern of smaller towns seemed to have an increase of crime rates during COVID while larger cities like Baltimore had a decrease in crime. 


### Task 8: Tell the Best Story

Write up a lede and nutgraf of a story with the _best_ story idea you've found in this assignment. The lede should be a single sentence that summarizes the most important information in the data. The nutgraf should be a paragraph that provides more context and explains why the data is important.

Violent crime dramatically increased in several smaller Maryland cities in 2020.

An analysis of 2019 and 2020 crime data across Maryland cities shows that crime trends weren’t consistent statewide, they really depended on the size and location of the city. Larger places like Frederick and Baltimore saw noticeable drops in property crime, in some cases by more than 25%. At the same time, several smaller towns reported sharp percentage increases in violent crime, but those jumps were often the result of just a few additional cases.



### Task 9: Save Your Work and Put it on GitHub

Save this notebook, then commit it to your local repository. Push the changes to your GitHub repository and submit the URL for this file in ELMS.

###Task 10:Percentiles
```{r}
md_cities_percentiles <- md_cities_rates |>
  mutate(percentile = ntile(violent2020, 100))
  md_cities_percentiles |>
  filter(percentile > 90) |>
  tally()
```  
  There are 10 cities above the 90th percentile. 
```{r}
md_cities_rates |>
  summarize(IQR_2020 = IQR(violent2020))
```  

In a way, my story from the previous assignment changes because the new analysis shows violent crime is not broadly rising across small Maryland cities but is concentrated in a relatively small group of places, that being said dramatic percentage increases in smaller towns can be considered misleading especially after viewing the percentiles.
