
---
title: "HW6_Spotify_Songs"
---

In this assignment, you will be working with a dataset of songs played on Spotify in 2020, and you'll be writing code to explore the range of danceability scores in the data, using standard deviation and histograms. You'll also write up your findings.

### Task 1: Load the tidyverse

```{r}
library(tidyverse)
```

### Task 2: Load the Spotify Songs dataset from https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/spotify_2020.csv

```{r}
spotify_songs <- read_csv("https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/spotify_2020.csv")
```

### Task 3: Calculate the mean and standard deviation of the danceability scores

Write code to calculate the mean and standard deviation of the danceability scores in the Spotify dataset. Display those.

```{r}
spotify_songs |> summarize(mean = mean(danceability), sd = sd(danceability))
```

### Task 4: Create a histogram of the danceability scores

Write code to create a histogram of the danceability scores in the Spotify dataset, based on the example we did in class.

```{r}
spotify_songs |> 
  ggplot() + 
  geom_histogram(aes(x= danceability), binwidth = 0.1) 
```

### Task 5: Add a line to the histogram for the mean danceability score

Take the same code and add a vertical line to the histogram for the mean danceability score, based on the example in class.

```{r}
spotify_songs |> 
  ggplot() + 
  geom_histogram(aes(x= danceability), binwidth = 0.1) +
  geom_vline(aes(xintercept = mean(danceability)), color = "red", linetype = "dashed", size = 1)
```


### Task 6: Add two lines to the histogram to show one standard deviation

Take the previous code and add two lines to the histogram for the 1st standard deviation below and above the mean score.

```{r}
spotify_songs |> 
  ggplot() + 
  geom_histogram(aes(x= danceability), binwidth = 0.1) +
  geom_vline(aes(xintercept = mean(danceability)), color = "red", linetype = "dashed", size = 1)+
geom_vline(aes(xintercept = mean(danceability) - sd(danceability)), color = "blue", linetype = "dashed", size = 1) +
  geom_vline(aes(xintercept = mean(danceability) + sd(danceability)), color = "blue", linetype = "dashed", size = 1)
```

### Task 7: Describe our results and pitch a story

Write a few sentences about the distribution of danceability scores in the Spotify dataset. Is it a normal distribution, or is it skewed in one direction? What does the mean and one standard distribution tell us about the songs in the dataset? Draft a short story pitch from this data.

The spotify song dataset is left skewed with an average danceability score of about 0.68. Since the data set is left-skewed, the results prove that music made for dancing is more popular. With 68% of the songs having a danceability score between 0.52 and 0.84, a main driver of streaming music is the ability to make people want to move. 

PITCH: Do Lyrics Even Matter Anymore?
In today’s music landscape, a catchy beat seems to dictate how popular a song is.Are listeners paying attention to lyrics or is the rhythm all that counts? This story would explore through both a data lens and through interviews 
the shift in music consumption. The story will dig into whether people really care about well written lyrics or just a catchy beat. 
### Task 8: Save Your Work and Put it on GitHub

Save this notebook, then commit it to your local repository. Push the changes to your GitHub repository and submit the URL for this file in ELMS.
